The origin of cosmic-ray electrons in cluster outskirts

### 2012-01-30 paper.tex
Method+updated f_z fig

% --- Anders's commands --- %
\newcommand{\ev}{\mathrm{eV}}
\newcommand{\gev}{\mathrm{GeV}}
\newcommand{\mev}{\mathrm{MeV}}
\newcommand{\tev}{\mathrm{TeV}}
\newcommand{\mpc}{m_{\mathrm{p}}c}
\newcommand{\mpcc}{m_{\mathrm{p}}c^2}
\newcommand{\p}{\mathrm{p}}
\newcommand{\Pp}{\mathrm{P}_\mathrm{p}}
\newcommand{\Pe}{\mathrm{P}_\mathrm{e}}
\newcommand{\pp}{p_\mathrm{p}}
\newcommand{\pe}{p_\mathrm{e}}
\newcommand{\br}{\bold{r}}
\newcommand{\e}{\mathrm{e}}
\newcommand{\ii}{\mathrm{i}}
\newcommand{\inj}{\mathrm{inj}}
\newcommand{\pd}{\partial}
\newcommand{\CR}{\mathrm{CR}}
\newcommand{\gas}{\mathrm{gas}}
\newcommand{\mr}{\mathrm}
\newcommand{\DM}{\mathrm{DM}}
\newcommand{\ISM}{\mathrm{ISM}}
\newcommand{\ICM}{\mathrm{ICM}}
\newcommand{\CMB}{\mathrm{CMB}}
\newcommand{\IC}{\mathrm{IC}}
\newcommand{\mecc}{m_\e c^2}
\newcommand{\mug}{\mathrm{\mu G}}
\newcommand{\rvir}{R_{200}}
\newcommand{\B}{\mathrm{B}}
\newcommand{\ma}{\mathrm{max}}
\newcommand{\low}{\mathrm{low}}
\newcommand{\gae}{\gamma_\mathrm{e}}
\newcommand{\gam}{\gamma}
\newcommand{\tend}{t_\rmn{end}}

@@

A recent spectacular example of radio gischt is CIZA J2242.8+5301, the 'sausage relic'  \citep{van-weeren10}, a large ($\sim 2$ Mpc long; located $\sim 1.5$ Mpc from the cluster center) double radio relic system. The post-shock radio spectral index was used to infer the particle spectral slope and hence the shock compression ratio and Mach number (${\cal M} \sim 4.6$), while the decrease in the spectral index (from $\sim 0.6$ to $\sim 2.0$ across the relic‚ narrow $\sim 55$ kpc width) toward the cluster center---spectral aging due to synchrotron and inverse Compton losses---was used to infer magnetic field strengths of $\sim 5 \, \mu$G. The strong ($\sim 50-60 \%$) polarization can be attributed to magnetic field frozen into the compressed ICM, which has been aligned parallel to the shock. The properties of this and similar systems are distinct from fossil radio plasma, which are smaller, have curved, steeper spectra (due to aging), and lobe-like morphology. The power-law spectral index, spectral gradient, and enormous extent clearly support a diffusive shock acceleration (DSA) origin.

@@

\begin{figure}[ht]
\begin{center}
%\resizebox{!}{5.5cm}{\includegraphics{./figures/ensslin.eps}}
%%\resizebox{!}{5.5cm}{\includegraphics{ensslin2.pdf}}
%\resizebox{!}{5.0cm}{\includegraphics{circles.pdf}}
%%\hspace{0.05cm}
%%%\vspace{-1cm} 
\end{center}
\caption{Cooling time of CR electrons (left panel) and CR protons (right panel) for typical conditions in the ICM. CRe can only survive for a Hubble time in the underdense outskirts of clusters; $\sim 100$MeV electrons should be the most long-lived. By contrast, CRp above 10 GeV can survive on cosmological timescales. From \citet{enslin11}. {\bf I stole this figure from Ensslin…should we make our own?}}
\label{fig:ensslin} % caption for the whole figure
\end{figure}

@@

\section{Relics}
\label{section:relics}
RELICS...  Collisionless cluster shocks are able to accelerate ions
and electrons in the high-energy tail of their Maxwellian distribution
functions through diffusive shock acceleration \citep[for reviews
  see][]{drury83, 1987PhR...154....1B,
  2001RPPh...64..429M}. Neglecting non-linear shock acceleration and
cosmic ray modified shock structure, then electrons and protons are
indistinguishable in the process of diffusive shock
acceleration. There are, however, difference in: (1) the maximum
energy of the steady state spectrum that depends on the details of the
shock, (2) acceleration efficiency that is related to the smaller
Larmor radius of the electrons that keeps the electrons from diffusing
back and forth over the discontinuity of the shock front.CHECK

\section{Simulations}
\label{sect:simulations}
Our simulations were performed in a $\Lambda$CDM universe using the
cosmological parameters: $\Omega_{\mr m}=\Omega_\DM + \Omega_{\mr
  b}=0.3, \, \Omega_{\mr b}=0.039, \, \Omega_\Lambda=0.7, \, h=0.7, \,
n_{\mr s}=1$, and $\sigma_8=0.9$. The total matter density in units of
the critical density of the universe $\rho_{\rm crit}$ is denoted by
$\Omega_{\mr m}$, the baryonic density by $\Omega_{\mr b}$, the DM
density by $\Omega_\DM$ and the cosmological constant today is denoted
by $\Omega_\Lambda$. The critical density, $\rho_{\rm crit}=3H_0/(8\pi
G)$, where the present day Hubble constant $H_0=100\, h \, \mr{km}\,
\mr{s}^{-1} \, \mr{Myr}^{-1}$. $n_{\mr s}$ represents the spectral
index of the primordial power-spectrum, and $\sigma_8$ denotes the
$rms$ linear mass fluctuation within a sphere of radius $8 \,h^{-1}
\mr{Mpc}$ extrapolated to $z = 0$. The simulations were carried out
with an updated and extended version of the distributed-memory
parallel TreeSPH code GADGET-2 \citep{2005MNRAS.364.1105S,
  2001NewA....6...79S}. Gravitational forces were computed using a
combination of particle-mesh and tree algorithms.  Hydrodynamic forces
are computed with a variant of the smoothed particle hydrodynamics
(SPH) algorithm that conserves energy and entropy where appropriate,
i.e. outside of shocked regions \citep{2002MNRAS.333..649S}.  Our
simulations follow the radiative cooling of the gas, star formation,
supernova feedback, and a photo-ionizing background \citep[details can
  be found in][]{2007MNRAS.378..385P}. We model the cosmic ray (CR)
physics in a self-consistent way \citep{pfrommer06,
  2007A&A...473...41E, 2008A&A...481...33J}.  We include the adiabatic
CR transport process such as compression and rarefaction, and a number
of physical source and sink terms which modify the cosmic ray pressure
of each CR population separately. The most important sources
considered for injection are, diffusive shock acceleration at
cosmological structure formation shocks and shock waves in supernova
remnants, while the primary sinks are thermalization by Coulomb
interactions, and catastrophic losses by hadronization. For
simplicity, in this paper we do not take into account CRs injected
into the inter-stellar medium from supernova remnants (see Pinzke, Oh,
and Pfrommer, in prep. for a discussion of this topic).

@@

10 & g676 & CC    & $1.3\times 10^{14}$ &   1.0~~ & 1.7  & 1.049 \\
11 & g914 & CC    & $1.2\times 10^{14}$ &   1.0~~ & 1.6  & 1.049 \\
12 & g1e  & M     & $9.1\times 10^{13}$ &  0.93   & 1.3  & 1.127 \\
13 & g8c  & M     & $8.5\times 10^{13}$ &  0.91   & 1.3  & 1.127 \\
14 & g8d  & PreM  & $7.8\times 10^{13}$ &  0.88   & 1.2  & 1.127 \\
\hline
\end{tabular}  \begin{quote}
 Notes:\\ (1) The dynamical state has been classified through a combined
 criterion invoking a merger tree study and the visual inspection of the X-ray
 brightness maps. The labels for the clusters are M--merger, PostM--post
 merger (slightly elongated X-ray contours, weak cool core (CC) region developing),
 PreM--pre-merger (sub-cluster already within the virial radius), CC--cool
 core cluster with extended cooling region (smooth X-ray profile).  (2) The
 virial mass and radius are related by $M_\Delta(z) = \frac{4}{3} \pi\,
 \Delta\, \rho_\rmn{crit}(z) R_\Delta^3 $, where $\Delta=200$ denotes a
 multiple of the critical overdensity $\rho_\rmn{crit}(z) = 3 H (z)^2/ (8\pi
 G)$.  (3) The virial temperature is defined by $kT_{\rm vir} = G M_{\rm vir} \,
 \mu\, m_\p / (2 R_{\rm vir})$, where $\mu$ denotes the mean molecular weight.
 (4) Time difference between output snapshots; in units of Myr for g72, and 
 remaining clusters show the ratio of the cosmological scalefactor between 
 two snapshots.
\end{quote}
\label{tab:cluster_sample}
\end{table}

\section{Method}
\label{sect:method}
In this section we outline the formalism that we use to derive the CR
electron spectra that builds up during the structure formation of
galaxy clusters, and the expected synchrotron emission from in galaxy
cluster that is undergoing a merger at low redshift.

Our scheme is summarized by:
\begin{itemize}
\item derive CR proton distribution function from each simulation output at different time $t$
\item calculate injected CR proton population between snapshots 
\item convert injected protons to injected electrons
\item evolve each injected CR electron population to a later time $\tend$ by
  accounting for losses (Coulomb, inverse Compton, and adiabatic)
\item add up the CR electron distribution for all SPH particles at time $\tend$
\item re-accelerate CR electron distribution function at time $\tend$
  using typical values for a merging shock in cluster outskirts
\item calculate the synchrotron radio emission from radio relic
\end{itemize}

\subsection{CR protons}
For simplicity we have attach a distribution function of CR protons to
each SPH fluid element. It is represented by a one-dimensional
isotropic distribution function \footnote{The three dimensional
  distibution function is given by $4\pi \int f(p)\, p^2\, \dd p$.}
given by
\begin{equation}
  \label{eq:f_p}
  f_\p(p_\p) = \frac{\dd^2 N_\p}{\dd p_\p\,\dd V} = 
  C\, p_\p^{-\alpha}\,\theta(p_\p-q)\,, 
\end{equation}
and $p_\p = P_\p/m_\p c$, where we have normalized the proton momentum
$P_\p$ with the proton mass $m_\p$. Here $q$ is the mass normalized
momentum cutoff, $\alpha$ the spectral index, and $C$ the
normalization of the distribution function in units of density.

\del{From our simulated galaxy clusters we derive for each snapshot at time
$t$ the injected CR proton distribution function $f_\p$. We select
particles in the outskirts of clusters where the gas densities are low
and hence the Coulomb cooling of the CR protons small. }

The CR proton distribution function from the simulations is
convinently parameterized in terms of adiabatic invariant momentum
cutoff $q_0$ and the adiabatic invariant Lagrangian amplitude of the
spectrum $\tilde{C}_0$. We convert the adiabatic invariant quantities
to physical quantities through
\begin{equation}
  \label{eq:q0}
  q=\left(\frac{\rho}{\rho_0}\right)^{\frac{1}{3}}q_0\,,
\end{equation}
and
\begin{equation}
  \label{eq:C0}
  \tilde{C}=\left(\frac{\rho}{\rho_0}\right)^{-\frac{\alpha-1}{3}}\tilde{C}_0\,,
\end{equation}
where the convenient unitless redefinition of the CR proton amplitude
is given by
\begin{equation}
  \label{eq:Ct}
  \tilde{C}=C\,m_\p/\rho\,.
\end{equation}

For each time $t$ we derive a CR proton distribution function $f_\p$
from the snapshots of our simulated galaxy clusters.

@@

The injected distribution function is calculated as the change in
CR normalization between a snapshot at time $t$ and an earlier time 
$t-\Delta t$:
\begin{eqnarray}
  \label{eq:finj}
f_\rmn{inj,p}(t,p_\p) &=& \Delta C(t)\,\frac{\rho}{m_\p}\,\rmn{AD(\rho)} \,p_\p^{-\alpha}\qquad\rmn{where}\\
  \Delta  C(t) &=& C(t) - C(t-\Delta t)\,
\end{eqnarray}
and the adiabatic cooling term $\rmn{AD}(\rho)\equiv\left(\rho/\rho_0\right)^{-1/3}$.

We fix the time between each snapshot $\Delta t$ to $100$~Myrs, which
is smaller then the loss timescale of CR electrons in cluster
outskirts.

\del{The CR proton number density is given by
\begin{equation}
  \label{eq:ncr}
  n_{\CR} = \int_0^\infty \dd p_\p\, f_\p(p_\p) =
  \frac{C\, q^{1-\alpha}}{\alpha-1}\,,
\end{equation}
provided $\alpha >1$. The kinetic energy density of the CR
proton population is
\begin{eqnarray}
  \label{eq:epscr}
  \eps_\CR &=& \int_0^\infty  \dd p_\p\, f_\p(p_\p) \,T_{\rmn
    p}(p_\p)=\frac{C\, m_\p\,c^2}{\alpha-1} \, \times
  \nonumber \\
  && \left[\frac{1}{2}
    \, \B_{\frac{1}{1+q^2}} \left(
    \frac{\alpha-2}{2},\frac{3-\alpha}{2}\right) + q^{1-\alpha}
    \left(\sqrt{1+q^2}-1 \right) \right] \,,
\end{eqnarray}
where $T_\p(p_\p) = (\sqrt{1+p_\p^2} -1)\, m_\p\,c^2$ is the kinetic energy
of a CR proton. $\B_x(a,b)$ denotes the incomplete Beta-function, and
$\alpha>2$ is assumed.}

\subsection{CR electrons}
In this section we derive the CR electron distribution function by
rescaling the injected CR proton spectrum to account for the different
acceleration efficiencies as well as the shift in momentum due to the
factor $\sim 2000$ difference in mass. In addition we model the
Coulomb and radiative losses of the CR electrons.

In a hot plasma the temperatures of electrons and protons (ions) are
equal. Under this constraint we derive the relationship between CR
electron momentum and CR proton momentum given by
%\begin{eqnarray}
%  \label{eq:pe_pp}
%\dd \Pe = \dd \Pp \frac{\sqrt{\Pe^2+2\left(m_\e c\right)^2-2m_\e m_\p\,c^2
%+2\left(m_\p-m_\e\right)c\sqrt{\left(m_\e c\right)^2+\Pe^2}}}{\Pe+\left(m_\p-m_\e\right) c \frac{\Pe}{\sqrt{\left(m_\e c\right)^2+\Pe^2}}}
%\end{eqnarray}
\begin{eqnarray}
  \label{eq:pe_pp}
\dd \pe &=& \dd \pp \,g(\pe) =  \dd \pp \,h(\pp)\,,\qquad \rmn{where}\\
g(\pe) &=& X \,\frac{\left[\pe^2-2\left(X-1\right)\left(1-\sqrt{1+\pe^2}\right)\right]^{0.5}}{\displaystyle\pe+\left(X-1\right)\frac{\pe}{\sqrt{1+\pe^2}}}\,,\qquad \rmn{and}\\
h(\pp) &=& X \,\frac{\displaystyle\pp-\left(1-Y\right)\frac{\pp}{\sqrt{1+\pp^2}}}{\left[\pp^2+2\left(1-Y\right)\left(1-\sqrt{1+\pp^2}\right)\right]^{0.5}}\,.\\
\end{eqnarray}
Note that $\pe = \Pe/m_\e c$ and $\pp = \Pp/m_\p c$, where we have
normalized the electron momentum $\Pe$ and the proton momentum $\Pp$
with the electron mass $m_\e$ and proton mass $m_\p$, respectively. We
have introduced the mass ratios $X=m_\p/m_\e$ and $Y=m_\e/m_\p$. Also
note that $\dd \Pe\rightarrow \dd \Pp$ in the asymptotic limit where
$\pp\gg 1$, and that $\dd \Pe\rightarrow \sqrt{m_\e / m_\p}\,\dd \Pp$ for
$\pp\ll 1$.

The injected CR electron distribution function at each time $t$ is
given by
\begin{eqnarray}
  \label{eq:f_inj_e}
  f_\rmn{inj,e}(p_\e) &=&  \frac{\dd^2 N_\rmn{inj,e}}{\dd V \dd p_\e}(p_\e) = 
\frac{\dd^2 N_\rmn{inj,e}}{\dd V \dd p_\p}(p_\e)\frac{1}{g(p_\e)} \nonumber\\
&=& \frac{\dd^2 N_\rmn{inj,p}}{\dd V \dd p_\p}(p_\p)
\left(\frac{p_\e}{p_\p}\right)^{-\alpha}\frac{\eta_\rmn{max,e}}{\eta_\rmn{max,p}}\frac{1}{g(p_\e)} = 
\Delta C\,p_\e^{-\alpha}\frac{\eta_\rmn{max,e}}{\eta_\rmn{max,p}}\frac{1}{g(p_\e)}\,.
 \nonumber\\
&&
\end{eqnarray}
We use that maximal 50\% of the energy available in a shock is
injected into CR protons ($\eta_\rmn{max,p}$) and a factor 10 smaller
efficiency of 5\% for the CR electrons ($\eta_\rmn{max,e}$).

>>>>

% --- Anders's commands --- %
\newcommand{\ev}{\mathrm{eV}}
\newcommand{\gev}{\mathrm{GeV}}
\newcommand{\mev}{\mathrm{MeV}}
\newcommand{\tev}{\mathrm{TeV}}
\newcommand{\mpc}{m_{\mathrm{p}}c}
\newcommand{\mpcc}{m_{\mathrm{p}}c^2}
\newcommand{\p}{\mathrm{p}}
\newcommand{\Pp}{\mathrm{P}_\mathrm{p}}
\newcommand{\Pe}{\mathrm{P}_\mathrm{e}}
\newcommand{\pp}{p_\mathrm{p}}
\newcommand{\pe}{p_\mathrm{e}}
\newcommand{\br}{\bold{r}}
\newcommand{\e}{\mathrm{e}}
\newcommand{\ii}{\mathrm{i}}
\newcommand{\inj}{\mathrm{inj}}
\newcommand{\pd}{\partial}
\newcommand{\CR}{\mathrm{CR}}
\newcommand{\gas}{\mathrm{gas}}
\newcommand{\mr}{\mathrm}
\newcommand{\DM}{\mathrm{DM}}
\newcommand{\ISM}{\mathrm{ISM}}
\newcommand{\ICM}{\mathrm{ICM}}
\newcommand{\CMB}{\mathrm{CMB}}
\newcommand{\IC}{\mathrm{IC}}
\newcommand{\mecc}{m_\e c^2}
\newcommand{\mug}{\mathrm{\mu G}}
\newcommand{\rvir}{R_{200}}
\newcommand{\B}{\mathrm{B}}
\newcommand{\ma}{\mathrm{max}}
\newcommand{\low}{\mathrm{low}}
\newcommand{\gae}{\gamma_\mathrm{e}}
\newcommand{\gap}{\gamma_\mathrm{p}}
\newcommand{\gam}{\gamma}
\newcommand{\tend}{t_f}

@@

A recent spectacular example of radio gischt is CIZA J2242.8+5301, the 'sausage relic'  \citep{van-weeren10}, a large ($\sim 2$ Mpc long; located $\sim 1.5$ Mpc from the cluster center) double radio relic system. The post-shock radio spectral index was used to infer the particle spectral slope and hence the shock compression ratio and Mach number (${\cal M} \sim 4.6$), while the decrease in the spectral index (from $\sim 0.6$ to $\sim 2.0$ across the relicâ narrow $\sim 55$ kpc width) toward the cluster center---spectral aging due to synchrotron and inverse Compton losses---was used to infer magnetic field strengths of $\sim 5 \, \mu$G. The strong ($\sim 50-60 \%$) polarization can be attributed to magnetic field frozen into the compressed ICM, which has been aligned parallel to the shock. The properties of this and similar systems are distinct from fossil radio plasma, which are smaller, have curved, steeper spectra (due to aging), and lobe-like morphology. The power-law spectral index, spectral gradient, and enormous extent clearly support a diffusive shock acceleration (DSA) origin.

@@

\begin{figure}[ht]
\begin{center}
%\resizebox{!}{5.5cm}{\includegraphics{./figures/ensslin.eps}}
%%\resizebox{!}{5.5cm}{\includegraphics{ensslin2.pdf}}
%\resizebox{!}{5.0cm}{\includegraphics{circles.pdf}}
%%\hspace{0.05cm}
%%%\vspace{-1cm} 
\end{center}
\caption{Cooling time of CR electrons (left panel) and CR protons (right panel) for typical conditions in the ICM. CRe can only survive for a Hubble time in the underdense outskirts of clusters; $\sim 100$MeV electrons should be the most long-lived. By contrast, CRp above 10 GeV can survive on cosmological timescales. From \citet{enslin11}. {\bf I stole this figure from EnsslinÉshould we make our own?}}
\label{fig:ensslin} % caption for the whole figure
\end{figure}

@@

\section{Simulations}
\label{sect:simulations}
Our simulations were performed in a $\Lambda$CDM universe using the
cosmological parameters: $\Omega_{\mr m}=\Omega_\DM + \Omega_{\mr
  b}=0.3, \, \Omega_{\mr b}=0.039, \, \Omega_\Lambda=0.7, \, h=0.7, \,
n_{\mr s}=1$, and $\sigma_8=0.9$. The total matter density in units of
the critical density of the universe $\rho_{\rm crit}$ is denoted by
$\Omega_{\mr m}$, the baryonic density by $\Omega_{\mr b}$, the DM
density by $\Omega_\DM$ and the cosmological constant today is denoted
by $\Omega_\Lambda$. The critical density, $\rho_{\rm crit}=3H_0/(8\pi
G)$, where the present day Hubble constant $H_0=100\, h \, \mr{km}\,
\mr{s}^{-1} \, \mr{Myr}^{-1}$. $n_{\mr s}$ represents the spectral
index of the primordial power-spectrum, and $\sigma_8$ denotes the
$rms$ linear mass fluctuation within a sphere of radius $8 \,h^{-1}
\mr{Mpc}$ extrapolated to $z = 0$. The simulations were carried out
with an updated and extended version of the distributed-memory
parallel TreeSPH code GADGET-2 \citep{2005MNRAS.364.1105S,
  2001NewA....6...79S}. Gravitational forces were computed using a
combination of particle-mesh and tree algorithms.  Hydrodynamic forces
are computed with a variant of the smoothed particle hydrodynamics
(SPH) algorithm that conserves energy and entropy where appropriate,
i.e. outside of shocked regions \citep{2002MNRAS.333..649S}.  Our
simulations follow the radiative cooling of the gas, star formation,
supernova feedback, and a photo-ionizing background \citep[details can
  be found in][]{2007MNRAS.378..385P}. We model the cosmic ray (CR)
physics in a self-consistent way \citep{pfrommer06,
  2007A&A...473...41E, 2008A&A...481...33J} and attach a CR proton
distribution function to each SPH fluid element. We include the
adiabatic CR transport process such as compression and rarefaction,
and a number of physical source and sink terms which modify the cosmic
ray pressure of each CR population separately. The most important
sources considered for injection are, diffusive shock acceleration at
cosmological structure formation shocks and shock waves in supernova
remnants, while the primary sinks are thermalization by Coulomb
interactions, and catastrophic losses by hadronization. For
simplicity, in this paper we do not take into account CRs injected
into the inter-stellar medium from supernova remnants (see Pinzke, Oh,
and Pfrommer, in prep. for a discussion of this topic).

@@

10 & g676 & CC    & $1.3\times 10^{14}$ &   1.0~~ & 1.7  & 1.049 \\
11 & g914 & CC    & $1.2\times 10^{14}$ &   1.0~~ & 1.6  & 1.049 \\
12 & g1e  & M     & $9.1\times 10^{13}$ &  0.93   & 1.3  & 1.127 \\
13 & g8c  & M     & $8.5\times 10^{13}$ &  0.91   & 1.3  & 1.127 \\
14 & g8d  & PreM  & $7.8\times 10^{13}$ &  0.88   & 1.2  & 1.127 \\
\hline
\end{tabular}  \begin{quote}
 Notes:\\ (1) The dynamical state has been classified through a combined
 criterion invoking a merger tree study and the visual inspection of the X-ray
 brightness maps. The labels for the clusters are M--merger, PostM--post
 merger (slightly elongated X-ray contours, weak cool core (CC) region developing),
 PreM--pre-merger (sub-cluster already within the virial radius), CC--cool
 core cluster with extended cooling region (smooth X-ray profile).  (2) The
 virial mass and radius are related by $M_\Delta(z) = \frac{4}{3} \pi\,
 \Delta\, \rho_\rmn{crit}(z) R_\Delta^3 $, where $\Delta=200$ denotes a
 multiple of the critical overdensity $\rho_\rmn{crit}(z) = 3 H (z)^2/ (8\pi
 G)$.  (3) The virial temperature is defined by $kT_{\rm vir} = G M_{\rm vir} \,
 \mu\, m_\p / (2 R_{\rm vir})$, where $\mu$ denotes the mean molecular weight.
 (4) Time difference between output snapshots; in units of Myr for g72, and 
 remaining clusters show the ratio of the cosmological scale factor between 
 two snapshots.
\end{quote}
\label{tab:cluster_sample}
\end{table}

\section{Method}
\label{sect:method}
In this section we explain both the method that we use to derive the
CR electron distribution function that builds up during the structure
formation process of galaxy clusters, and the formalism for
calculating synchrotron emission from in galaxy cluster that is
undergoing a merger at low redshifts.

Our scheme is summarized in decending order by the following bullet points:
\begin{itemize}
\item derive CR proton distribution function from each simulation output at different time $t$
\item calculate injected CR proton population between snapshots 
\item convert injected protons to injected electrons
\item evolve each injected CR electron population to a later time $\tend$ by
  accounting for losses (Coulomb, inverse Compton, and adiabatic)
\item add up the CR electron distribution for all SPH particles at time $\tend$
\item re-accelerate CR electron distribution function at time $\tend$
  using typical values for a merging shock in cluster outskirts
\item calculate the synchrotron radio emission from radio relic
\end{itemize}

\subsection{CR protons}
The CR protons in the simulations are represented by a one-dimensional
isotropic distribution function \footnote{The three dimensional
  distribution function is given by $4\pi \int f(\gam)\, \gam^2\, \dd \gam$.}
given by
\begin{equation}
  \label{eq:f_p}
  f_\p(\gap) = \frac{\dd^2 N_\p}{\dd \gap\,\dd V} = 
  C\, \gap^{-\alpha}\,\theta(\gap-q)\,, 
\end{equation}
and $\gap \approx P_\p/m_\p c$, where we have normalized the proton momentum
$P_\p$ with the proton mass $m_\p$. Here $q$ is the mass normalized
momentum cutoff, $\alpha$ the spectral index, and $C$ the
normalization of the distribution function in units of density.

To improve the numerical efficiency of the simulations the CR proton
distribution function is parametrized in terms of adiabatic invariant
momentum cutoff $q_0$ and the adiabatic invariant Lagrangian amplitude
of the spectrum $\tilde{C}_0$. We convert the adiabatic invariant
quantities to physical quantities through
\begin{equation}
  \label{eq:q0}
  q=\left(\frac{\rho}{\rho_0}\right)^{\frac{1}{3}}q_0\,,
\end{equation}
and
\begin{equation}
  \label{eq:C0}
  \tilde{C}=\left(\frac{\rho}{\rho_0}\right)^{-\frac{\alpha-1}{3}}\tilde{C}_0\,,
\end{equation}
where the convenient unitless redefinition of the CR proton amplitude
is given by
\begin{equation}
  \label{eq:Ct}
  \tilde{C}=C\,m_\p/\rho\,.
\end{equation}

@@

The CR electron distribution function can not be derived directly
from the CR protons because of the different timescales of
losses. Instead we first need to derive the injected CR proton
distribution function between each snapshot.  The injected distribution
function is calculated as the change in CR normalization between a
snapshot at time $t$ and an earlier time $t-\Delta t$:
\begin{eqnarray}
  \label{eq:finj}
f_\rmn{inj,p}(t,\gap) &=& \Delta C(t)\,\frac{\rho}{m_\p}\,\rmn{X(\rho)} \,
\gap^{-\alpha}\qquad\rmn{where}\\
  \Delta  C(t) &=& C(t) - C(t-\Delta t)\,
\end{eqnarray}
and the adiabatic cooling term
$\rmn{X}(\rho)\equiv\left(\rho/\rho_0\right)^{-1/3}$. The time between
snapshots is denoted by $\Delta t$ and is shown in
Table~\ref{tab:cluster_sample} for each simulated cluster. Note that
for most clusters we can not resolve the cooling timescale of CR
electrons, and hence we over predict the abundance of electrons for
low and high momenta (see Figure~\ref{fig:e_spec_tests} for more
details) for some clusters. Although, in the region between $\gamma ~
(10-10^3)$ is small enough to resolve the cooling processes.

\subsection{CR electrons}
The main difference between the CR electrons and CR protons comes from
the less efficent acceleration of the electrons due to the smaller
gyroradius, and typically the larger losses. We derive the CR electron
distribution function by renormalizing the injected CR proton spectrum to
account for the different acceleration efficiencies as well as the
shift in momentum due to the factor $\sim 2000$ difference in mass. In
addition we model the Coulomb and radiative losses of the CR
electrons.

The injected CR electron distribution function at each time $t$ is
given by
\begin{eqnarray}
  \label{eq:f_inj_e}
  f_\rmn{inj,e}(\gae,t) &\equiv&  
\frac{\dd^2 N_\rmn{inj,e}}{\dd V \dd \gap}(\gae,t) = 
\frac{\dd^2 N_\rmn{inj,e}}{\dd V \dd \gae}(\gae,t)\,Y \nonumber\\
&=& \frac{\dd^2 N_\rmn{inj,p}}{\dd V \dd \gap}(\gap,t)
\left(\frac{\gae}{\gap}\right)^{-\alpha}\frac{\eta_\rmn{max,e}}{\eta_\rmn{max,p}}Y = 
\Delta C(t)\,\gae^{-\alpha}\frac{\eta_\rmn{max,e}}{\eta_\rmn{max,p}}Y\,.
 \nonumber\\
&& \\
\rmn{where}&& Y=\frac{m_\p}{m_\e}
\end{eqnarray}
We use that maximal 50\% of the energy available in a shock is
injected into CR protons ($\eta_\rmn{max,p}$) and a factor 10 smaller
efficiency of 5\% for the CR electrons ($\eta_\rmn{max,e}$). Note that
for readability we now drop the electron index $\e$ on the lorentz
factor $\gae$ for the electrons.

The CR electrons cool through inverse Compton (IC) emission and
Coulomb losses on timescales that are relative short compared to the
dynamical timescale of a cluster. We model these looses analytically
by instantaneously injecting a power-law of electrons at time $t_i$
and evolving it to a later time $t_f$ (for further details see
\citep{1999ApJ...520..529S}). The loss of energy for each particle is
described by
\begin{equation}
\label{eq:d_cool}
\frac{\dd \gam}{\dd t} = - b(\gam,t)\,,
\end{equation}
where the loss function $b(\gam,t)$ is dominated by Coulomb and IC
losses for the energies and gas densities relevant in this work. The
Coulomb losses are given by
\begin{eqnarray}
  \label{eq:b_C}
  b_\rmn{C}(\gam) &=& b_\rmn{C} \gam^2 
  = \frac{3\,\sigma_\rmn{T}\,n_\rmn{el}\,c}{2}
  \left[\ln\left(\frac{m_\e c^2 \sqrt{\gam-1}}{\hbar\,\omega_\rmn{plasma}}\right)\right.\nonumber \\
    &-&\ln(2)\left(\frac{1}{2}+\frac{1}{\gam}\right)+\frac{1}{2}+
    \left.\left(\frac{\gam-1}{4\gam}\right)^2\right]\,.
\end{eqnarray}
Here $\omega_\rmn{plasma} = \sqrt{4\pi e^2 n_\e / m_\e}$ is the plasma
frequency, and $n_\e$ is the number density of free electrons. Since
the Coulomb losses for relativistic electrons is almost independent of
energy (logarithmic), we approximate that $b_\rmn{C}(\gam,t)\approx
b_\rmn{C}(t)$ \footnote{We use a constant energy $\gamma=10^2$ to
  model the very weak dependence of energy, and note that the
  particular value of $\gamma$ does not matter as long as
  $\gamma\gg1$.}. We can now derive the shift in energy $\gam_i$ at
time $t_i$ to energy $\gam_f$ at time $t_f$ due to Coulomb cooling by
integrating Eqn.~(\ref{eq:d_cool}):
\begin{eqnarray}
  \label{eq:gamma_low}
  \gam_\low \equiv\int_{t_i}^{t_f}\dd t \, b_\rmn{C}(t) = 
  -\int_{\gam_i}^{\gam_f}\dd \gam' = -(\gam_f-\gam_i)\,.
\end{eqnarray}
Beacuse the snapshots are descrete in time, we approximate the low
momentum cut-off between time $t_i$ and the later time $t_f$ by
\begin{eqnarray}
  \label{eq:gamma_low_sum}
  \gam_\low \approx \sum_{j=i}^{f} \gam_\low(\Delta t_j) \approx
  \sum_{j=i}^{f} \Delta t_j \left[b_\rmn{C}(t_{j-1}) + b_\rmn{C}(t_j)\right]/2\,,
\end{eqnarray}
where $j$ denotes the summation over all snapshots between time $t_i$
and the later time $t_f$. Here we have approximated the low momentum
cut-off between two snapshots by the mean given by $\gam_\low(\Delta
t_j) = \Delta t_j \left[b_\rmn{C}(t_{j-1})+ b_\rmn{C}(t_j)\right]/2$,
where $\Delta t_j = t_j - t_{j-1}$.

We now continue with the IC losses that are given by
\begin{equation}
  \label{eq:b_IC}
  b_\IC(\gam,z) = b_\IC \gam^2 (1+z)^4
  = \frac{4}{3}\frac{\sigma_\rmn{T}}{m_\e c} U_\rmn{cmb} \gam^2\,.
\end{equation}
Here $\sigma_\rmn{T}= 8\pi e^4/3(m_\e c^2)^2$ is the Thomson cross
section and $U_\rmn{cmb}$ is the energy density of the CMB at redshift
$z=0$. Similarly to the Coulomb cooling we derive the evolution of energy 
of a particle subject to IC losses through
\begin{equation}
  \label{eq:b_IC_evolu}
\frac{\dd \gam}{\gam^2} = -b_\IC (1+z)^4\dd t\,.
\end{equation}
After a time $(t_f-t_i)$ has elapsed, all the electrons with
energies above $\gam_\ma$ have thermalized, where we derive $\gam_\ma$
through integrating Eqn.~(\ref{eq:b_IC_evolu}) from injected time
$z_i=z(t_i)$ to a later time $z_f=z(t_f)$:
\begin{equation}
  \label{eq:b_max}
   \frac{1}{\gam_\ma} \equiv \frac{1}{\gam_f} - \frac{1}{\gam_i} = 
   \frac{b_\IC}{H_0}\left[\Lambda(z_f)-\Lambda(z_i)\right]\,.
\end{equation}
Here $\Lambda(z) \approx z + 1.23\,z^2 + 0.50\,z^3 - 0.14\,z^4 -
0.04\,z^5$ in a Lambda CDM universe and
$H_0=100\,h\,\rmn{km}\,\rmn{s}^{-1}\,\rmn{Mpc}^{-1}$ is the Hubble constant.
